Won’t you be my neighboR?

### 2019-05-16 msu-workshop-2019.Rmd
update

---
title: "Won’t you be my neighboR?"
subtitle: "An introduction to R for data science in education"
author: "Joshua Rosenberg"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

@@

1. Why learn R?
1. Getting started with R Studio
1. Computer science foundations
1. Processing data
1. Loading data
1. Visualizing data
1. Modeling data 
1. Learning and doing more with R
1. Questions and sundry tools (time permitting)

@@

* It is capable of carrying out basic and complex statistical analyses
* It is able to work with data small (*n* = 30!) and large (*n* = 100,000+) efficiently
* It is a programming language and so is quite flexible
* There is a great, inclusive community of users and developers (and teachers)
* It is cross-platform, open-source, and freely-available
* It can h
---

@@

# Running Functions from Packages

Once you have loaded the package in your session, you can run the functions that are contained within that package.

To find a list of all those functions, you can run this in the R Studio console:

```{r, eval = FALSE}
help(package = dplyr)
```

The documentation should tell you what the function does, what arguments (i.e., details) needed for it to successfully run, examples, and what the output should look like.

@@

We can also rename the variables at the same time we select them. I put these on separate lines so I could add the comment, but you could do this all in the same line, too. It does not make a difference in terms of how `select()` will work. We are using `head()` to only print a few rows.

```{r}
student_responses %>% # using dataframe student_responses
  select(student_efficacy = SCIEEFF,
         student_joy = JOYSCIE, 
         student_broad_interest = INTBRSCI,
         student_epistemic_beliefs = EPIST,
         student_instrumental_motivation = INSTSCIE,
  ) %>% 
  head(3)
```

@@

@@

@@

It is very flexible ... but can also be a bit challenging to use, at first.

@@

```{r, warning = FALSE}
ggplot(student_mot_vars,
       aes(x = student_efficacy, y = student_broad_interest)) +
  geom_point()
```

@@

```{r, warning = FALSE}
ggplot(student_mot_vars,
       aes(x = student_efficacy, y = student_broad_interest)) +
  geom_point() +
  geom_smooth(method = "lm")
```

@@

```{r, warning = FALSE}
ggplot(student_mot_vars,
       aes(x = student_efficacy, y = student_broad_interest)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() +
  xlab("Student Efficacy") +
  ylab("Student Broad Interest")
```

```{r, warning = FALSE}
ggplot(student_mot_vars,
       aes(x = student_efficacy, y = student_broad_interest)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() +
  xlab("Student Efficacy") +
  ylab("Student Broad Interest")
```

@@

@@

# Latent variable models

@@

class: inverse, center, middle
# Questions, other

@@

# R Markdown

@@

# APA Tables

@@

# Papaja

# Redoc

# Learning and doing more with R

# Following up

>>>>

---
title: "Won’t you be my neighboR?"
subtitle: "An introduction to R for data science in education"
author: "Joshua Rosenberg and Emily Bovee"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

@@

1. Why learn R?
1. Getting started with R Studio
1. Computer science foundations
1. Processing data
1. Loading data
1. Visualizing data
1. Modeling data 
1. Learning and doing more with R
1. Questions

@@

* It is capable of carrying out basic and complex statistical analyses
* It is able to work with data small (*n* = 30!) and large (*n* = 100,000+) efficiently
* It is a programming language and so is quite flexible
* There is a great, inclusive community of users and developers (and teachers)
* It is cross-platform, open-source, and freely-available

@@

@@

We can also rename the variables at the same time we select them. Let's save the results to `student_mot_vars`.

```{r}
student_mot_vars <- student_responses %>% 
  select(student_efficacy = SCIEEFF,
         student_joy = JOYSCIE, 
         student_broad_interest = INTBRSCI,
         student_epistemic_beliefs = EPIST,
         student_instrumental_motivation = INSTSCIE,
  )

student_mot_vars %>% 
  head(3)
```

@@

*Note. Check out [R Studio Cloud](https://rstudio.cloud/) if you're having any installation trouble!*

@@

# The grammar of graphics

@@

@@

Notice the three parts of all `ggplot2` graphs: a) the data (`student_mot_vars`), b) an `aes()`thetic mapping, and c) a `geom`:

```{r, warning = FALSE, out.width = "450px", echo = FALSE, fig.align = "center"}
ggplot(student_mot_vars,
       aes(x = student_efficacy, y = student_broad_interest)) +
  geom_point()
```

@@

```{r, warning = FALSE, out.width = "450px", fig.align = "center"}
ggplot(student_mot_vars,
       aes(x = student_efficacy, y = student_broad_interest)) +
  geom_point() +
  geom_smooth(method = "lm")
```

@@

```{r, warning = FALSE, out.width = "450px", fig.align = "center"}
ggplot(student_mot_vars,
       aes(x = student_efficacy, y = student_broad_interest)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() +
  xlab("Student Efficacy") +
  ylab("Student Broad Interest")
```

@@

The `lm()` function is a very helpful general purpose function for linear models. We'll use the `student_mot_vars` data frame we created earlier.

```{r}
lm(student_epistemic_beliefs ~ student_broad_interest, data = student_mot_vars)
```

Let's save the results back to an object, `m1`.

```{r}
m1 <- lm(student_epistemic_beliefs ~ student_broad_interest, data = student_mot_vars)
```

We can then run a summary on the output.

```{r}
summary(m1)
```

You can then build up a more complex model, i.e.:

```{r, eval = FALSE}
m2 <- lm(student_epistemic_beliefs ~ student_broad_interest + student_joy + student_broad_interest:student_joy, data = student_mot_vars)
```

You can then run `summary()` to view the results.

@@

```{r, eval = TRUE}
library(lme4)
m2_mlm <- lmer(student_epistemic_beliefs ~ student_broad_interest + student_joy + student_broad_interest:student_joy + (1|teacher_id), data = student_mot_vars)
m2_mlm
```
---

@@

class: inverse, center, middle
# Learning and doing more with R

@@

# Resources

  - [Doing data science with R](https://r4ds.had.co.nz/) by Wickham and
    Grolemund (2017)
  - [Big magic with R: Creating learning beyond
    fear](https://speakerdeck.com/apreshill/big-magic-with-r-creative-learning-beyond-fear)
    by Hill (2017)
  - [Data science in education](https://github.com/data-edu/data-science-in-education)

@@

# Courses

  - [\#r4ds](https://medium.com/@kierisi/r4ds-the-next-iteration-d51e0a1b0b82)
    (see a talk at rstudio::conf()
    [here](https://resources.rstudio.com/rstudio-conf-2019/r4ds-online-learning-community-improvements-to-self-taught-data-science-and-the-critical-need-for-diversity-equity-and-inclusion-in-data-science-education)
    by Maegan (2019))
  - [Data science for social scientists](http://datascience.tntlab.org/) by
    Landers (2019)
  - [University of Oregon Data Science Specialization for the College of
    Education](https://github.com/uo-datasci-specialization) by Anderson (2019)

@@

class: inverse, center, middle
# Questions?

The repository for this workshop is [here](https://github.com/jrosen48/MSU-workshop-2019).
